### 2020-04-27 pd2.Rmd
Comparing imputation times cyyk

---
title: "The imputation methods big test"
author: "Jakub Kosterna, Dawid Przybyliński & Hanna Zdulska"
date: "22/04/2020"
output:
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

Choosing the best-suited imputation is the daily dilemma of every data scientist. Some of they believe the crux lies in the most advanced and sophisticated, the others trust the simplest of all possible.

In this document we will try to find one objectively the finest imputation method by testing five of popular ones on eight specially selected data sets prepared for these operations.

## Imputation functions

We'll look at the two simple imputation methods with the following short markings:

1. *I1* / *modeMedian* - missing data supplemented with medians from individual columns
2. *I2* / *removeRows* - all the rows with any nonexistent values removed

```{r imputation_basic_functions, cache = TRUE}
imputation_mode_median <- function(df){

  Mode <- function(x) {
    ux <- unique(x)
    ux[which.max(tabulate(match(x, ux)))]
  }

  for (i in 1L:length(df)){
    if (sum(is.na(df[,i])) > 0){
      if (mode(df[,i]) == 'character' | is.factor(df[,i])){
        to_imp <- Mode(df[,i])
        df[,i][is.na(df[,i])] <- to_imp
      }
      else{
        to_imp <- median(df[,i], na.rm = TRUE) 
        df[,i][is.na(df[,i])] <- to_imp
      }
    }
  }

  return(df)
}
imputation_remove_rows <- function(df){
  return (na.omit(df))
}
```

... and also three more advanced algorithms from popular libraries:

3. *I3* - *mice*
4. *I4* - *vim*
5. *I5* - *missForest*

```{r imputation_advanced_functions, message = FALSE, warning = FALSE, cache = TRUE}
library(mice)
imputation_fun_mice <- function(df){
  init <- mice(df, maxit=0) 
  meth <- init$method
  predM <- init$predictorMatrix
  imputed <- mice(df, method=meth, predictorMatrix=predM, m=5)
  completed <- complete(imputed)
  return(completed)
}
library(VIM)
imputation_fun_vim <- function(df){
  no_columns <- length(df)
  imputed <- kNN(df)
  imputed <- imputed[,1:no_columns]
  return(imputed)
}
library(missForest)
imputation_fun_missForest <- function(df){
  return(missForest(df)$ximp)
}
```

## Reading datasets

The eight data frames on which we will test these above were taken from **OpenML100 collection** and were corrected specifically for this research. They can be found under the following identifiers with the following names:

* 29 - *credit-approval* PÓKI CO NIE BANGLA
* 38 - *sick*
* 56 - ...
* 188 - *eucalyptus*
* 1590 - *adult*
* 6332 - *cylinder-bands*
* 23381 - *dresses-sales*
* 40536 - *SpeedDating*
* 41278 - *okcupid-stem*

Those above have been placed in individual directories identified by id in the prepared directory.

```{r read_dataset, cache = TRUE}
DFT_REPO_DATASET_DIR = './dependencies/datasets'
read_dataset <- function(openml_id, dataset_dir = DFT_REPO_DATASET_DIR){

  if (!dir.exists(dataset_dir)){
    stop(paste(dataset_dir, 'does not exist' ))
  }

  dir <- paste(dataset_dir, paste('openml_dataset', openml_id, sep = '_'), sep ='/')
  if (!dir.exists(dir)){
    stop(paste(dir, 'does not exist' ))
  }

  start_dir <- getwd()

  # set right dir to code.R to acually work - it depends on dirlocation to create json
  setwd(dir)
  # use new env to avoid trashing globalenv
  surogate_env <- new.env(parent = .BaseNamespaceEnv)
  attach(surogate_env)
  source("code.R",surogate_env)

  j <- jsonlite::read_json('./dataset.json')
  j$dataset <- surogate_env$dataset
  setwd(start_dir)

  return(j)
}
```

In order to read all the eight datasets we will use other function, which will return a dataframe containing all the important informations about the test matrices.

```{r read_all_datasets, cache = TRUE}
read_all_datasets <- function(dataset_dir = DFT_REPO_DATASET_DIR){

  if (!dir.exists(dataset_dir)){
    stop(paste(dataset_dir, 'does not exist'))
  }

  start_dir <- getwd()
  subdirs <- dir(dataset_dir)
  ids <- sapply(subdirs, function(dir){substr(dir, 16, nchar(dir))})
  datasets_combined <- lapply(ids, function(x){read_dataset(x, dataset_dir)})

  datasets_combined <- t(datasets_combined)
  return(unname(t(datasets_combined)))
}
```

## Metrics functions

In order to test the same test-train splits we'll use one random seed 1357 for all datasets.

```{r train_test_split, cache = TRUE}
set.seed(1357)
train_test_split <- function(dataset, train_size){
  smp_size <- floor(train_size * nrow(dataset))
  typeof(smp_size)

  train_ind <- sample(seq_len(nrow(dataset)), size = smp_size)

  train <- dataset[train_ind, ]
  test <- dataset[-train_ind, ]

  return (list(train, test))
}
```

For each machine learning model after every imputation we will get the confusion matrix and the values of four basic metrics:

* *accuracy* - $\frac{TP+TN}{TP+FP+FN+TN}$
* *precision* - $\frac{TP}{TP+FP}$
* *recall* - $\frac{TP}{TP+FN}$
* *f1* - $2*\frac{Recall * Precision}{Recall + Precision}$

```{r metrics, cache = TRUE}
get_confusion_matrix <- function(test, pred){
  return (table(Truth = test, Prediction = pred))
}
confusion_matrix_values <- function(confusion_matrix){
  TP <- confusion_matrix[2,2]
  TN <- confusion_matrix[1,1]
  FP <- confusion_matrix[1,2]
  FN <- confusion_matrix[2,1]
  return (c(TP, TN, FP, FN))
}
accuracy <- function(confusion_matrix){
  conf_matrix <- confusion_matrix_values(confusion_matrix)
  return((conf_matrix[1] + conf_matrix[2]) / (conf_matrix[1] + conf_matrix[2] + conf_matrix[3] + conf_matrix[4]))
}
precision <- function(confusion_matrix){
  conf_matrix <- confusion_matrix_values(confusion_matrix)
  return(conf_matrix[1]/ (conf_matrix[1] + conf_matrix[3]))
}
recall <- function(confusion_matrix){
  conf_matrix <- confusion_matrix_values(confusion_matrix)
  return(conf_matrix[1] / (conf_matrix[1] + conf_matrix[4]))
}
f1 <- function(confusion_matrix){
  conf_matrix <- confusion_matrix_values(confusion_matrix)
  rec <- recall(confusion_matrix)
  prec <- precision(confusion_matrix)
  return(2 * (rec * prec) / (rec + prec))
}
```

## Imputation results

```{r get_result_function, warning = FALSE, message = FALSE, cache = TRUE}
library(rpart)
get_result <- function(dataset_list, imputation_fun){

dataset <- dataset_list$dataset
name_of_target <- dataset_list$target
# imputation
imputation_start = Sys.time() # start to measure time
imputated_dataset <- imputation_fun(dataset) 
imputation_stop = Sys.time() # end measuring time
# train test split
train_test <- train_test_split(imputated_dataset, 0.8)
train <- as.data.table(train_test[1])
test <- as.data.table(train_test[2])
# modelling
vars <- colnames(dataset)[colnames(dataset)!=name_of_target]
my_formula <- as.formula(paste(name_of_target, paste(vars, collapse=" + "), sep=" ~ "))
modelling_start = Sys.time() # start to measure time
tree_model <- rpart(formula = my_formula, data = train,
                    method = "class", control = rpart.control(cp = 0))
y_pred <- as.data.frame(predict(tree_model, test, type = "class"))
modelling_stop = Sys.time() # end measuring time
# calculating metrics
confusion_matrix <- get_confusion_matrix(test[[name_of_target]], y_pred[,1])
accuracy_v <- accuracy(confusion_matrix)
precision_v <- precision(confusion_matrix)
recall_v <- recall(confusion_matrix)
f1_v <- f1(confusion_matrix)
classification_report <- data.frame(accuracy_v, precision_v,
                                    recall_v, f1_v)
colnames(classification_report) <- c("accuracy", "precision",
                                     "recall", "f1")
dataset_list$dataset <- NULL
# in future maybe return all dataset_list ?
# for now stick with readability
imp_method_name <- deparse(substitute(imputation_fun))
return(list( dataset_id = dataset_list$id, 
             imp_method = imp_method_name,
             confusion_matrix = confusion_matrix,
             classification_report = classification_report,
             imputation_time = imputation_stop - imputation_start,
             modelling_time = modelling_stop - modelling_start))
}
data_all <- read_all_datasets()
# imputations and targets preparation
imputations <- list(imputation_fun_vim, imputation_fun_missForest,
                  imputation_remove_rows, imputation_mode_median, imputation_fun_mice)
targets <- lapply(data_all, function(d){d$target})
```

Here are the results:

```{r warning=FALSE, cache = TRUE}
  library(knitr)
  results <- rep(list(0), 5)
  results[[1]] <- readRDS('./part_results/res1_new.rds')
  results[[2]] <- readRDS('./part_results/res2_new.rds')
  results[[3]] <- readRDS('./part_results/res3_new.rds')
  results[[4]] <- readRDS('./part_results/res4_new.rds')
  results[[5]] <- readRDS('./part_results/res5_new.rds')

  result_table <- function(ds_it){
    d <- data.frame(matrix(ncol = 7, nrow = 0))
    colnames(d) <- c("imputation", "accuracy", "precision", "recall", "f1", "imp_time", "mod_time")
    for (i in 1:length(imputations)){
      if (length(results[[i]][[ds_it]]) != 1){ # length equal 1 means "ERROR" was generated in partial results 
        d[i,] <- c(results[[i]][[ds_it]]$imp_method,
               round(as.numeric(results[[i]][[ds_it]]$classification_report$accuracy),3),   
               round(as.numeric(results[[i]][[ds_it]]$classification_report$precision),3),
               round(as.numeric(results[[i]][[ds_it]]$classification_report$recall),3),
               round(as.numeric(results[[i]][[ds_it]]$classification_report$f1),3),
               round(as.numeric(results[[i]][[ds_it]]$imputation_time),3),
               # @TODO konwersja jednostek, as.numeric nie rozroznia jednostek czasu
               round(as.numeric(results[[i]][[ds_it]]$modelling_time),3)) 
      }
      else{ # for error input imputation name
        if (i==3){
          d[i,1] <- "imputation_fun_mice" 
        }
        if (i==4){
          d[i,1] <- "imputation_fun_missForest" 
        }
        if (i==5){
          d[i,1] <- "imputation_fun_vim" 
        }
      }
    }
    d
  }

  # change i bound to number of datasets 
  #for (i in 1:length(results[[1]])){ # won't use it, because the output looks bad
  #  kable(result_table(i), caption = paste("Dataset id: ", results[[2]][[i]]$dataset_id, sep = ""))
  #}

  kable(result_table(1), caption = paste("Dataset id: ", results[[2]][[1]]$dataset_id, sep = ""))
  kable(result_table(2), caption = paste("Dataset id: ", results[[2]][[2]]$dataset_id, sep = ""))
  kable(result_table(3), caption = paste("Dataset id: ", results[[2]][[3]]$dataset_id, sep = ""))
  kable(result_table(4), caption = paste("Dataset id: ", results[[2]][[4]]$dataset_id, sep = ""))
  kable(result_table(5), caption = paste("Dataset id: ", results[[2]][[5]]$dataset_id, sep = ""))
  kable(result_table(6), caption = paste("Dataset id: ", results[[2]][[6]]$dataset_id, sep = ""))
  kable(result_table(7), caption = paste("Dataset id: ", results[[2]][[7]]$dataset_id, sep = ""))
  kable(result_table(8), caption = paste("Dataset id: ", results[[2]][[8]]$dataset_id, sep = ""))
  kable(result_table(9), caption = paste("Dataset id: ", results[[2]][[9]]$dataset_id, sep = ""))
  kable(result_table(10), caption = paste("Dataset id: ", results[[2]][[10]]$dataset_id, sep = ""))
  kable(result_table(11), caption = paste("Dataset id: ", results[[2]][[11]]$dataset_id, sep = ""))
  kable(result_table(12), caption = paste("Dataset id: ", results[[2]][[12]]$dataset_id, sep = ""))
  kable(result_table(13), caption = paste("Dataset id: ", results[[2]][[13]]$dataset_id, sep = ""))
  kable(result_table(14), caption = paste("Dataset id: ", results[[2]][[14]]$dataset_id, sep = ""))
```

## Comparing imputation times

Let's build a dataframe containing information about imputation methods.

```{r imputation_times_counting, cache = TRUE}
ndatasets <- length(results[[1]])
nimputations <- length(imputations)
imputation_names <- c("remove", "median", "mice", "missForest", "VIM")
dataset_ids <- rep(NULL, ndatasets)
for(i in 1:ndatasets){
  dataset_ids[i] <- results[[2]][[i]]$dataset_id
}

imputation_times <- data.frame(rep(0, 14))
for (imputation_id in 1:nimputations){
  imp_id_times <- rep(NULL, ndatasets)

  for(dataset_id in 1:ndatasets){
    if("imputation_time" %in% names(results[[imputation_id]][[dataset_id]])){
      imp_time <- results[[imputation_id]][[dataset_id]]$imputation_time
      imp_time <- as.numeric(imp_time, units="secs")
      imp_id_times[dataset_id] <- imp_time
    }
  }
  imputation_times <- cbind(imputation_times, imp_id_times)
}

imputation_times[1] <- dataset_ids
colnames(imputation_times) <- c("dataset_id", imputation_names)
imputation_times$dataset_id <- as.factor(imputation_times$dataset_id)
knitr::kable(imputation_times)
```

Looks good! Now it's visualization time.

```{r imputation_times_plot, message = FALSE, warning = FALSE, cache = TRUE}
library(ggplot2)
# install.packages("reshape") # if not installed
library(reshape)

# in order to connect colors and shapes in legend see:
# https://stackoverflow.com/questions/12410908/combine-legends-for-color-and-shape-into-a-single-legend

colors <- c("remove" = "orange", "median" = "cyan", "mice" = "green",
            "missForest" = "red", "VIM" = "blue")
imptimes <- melt(imputation_times)

require(scales)
ggplot(data=imptimes, aes(x = dataset_id, y = value, color = variable, shape = variable)) + geom_point() + 
#ggplot(data = imputation_times, aes(x = dataset_id, y )) +
  #geom_point(aes(y = remove), color = colors[1], size = 2, shape = 18) +
  #geom_point(aes(y = median), color = colors[2], size = 2, shape = 16) +
  #geom_point(aes(y = mice), color = colors[3], size = 2, shape = 15) +
  #geom_point(aes(y = missForest), color = colors[4], size = 2, shape = 17) +
  #geom_point(aes(y = VIM), color = colors[5], size = 2, shape = 19) +
  ggtitle("Datasets' imputations times") +
  theme_gray() +
  scale_y_continuous(trans = log2_trans(),
    breaks = trans_breaks("log2", function(x) 2^x),
    labels = trans_format("log2", math_format(2^.x))) +
  labs(x = "Dataset Id",
       y = "Imputation time [in seconds]",
       color = "Legend") +
  scale_color_manual(name = "imputation name", values = colors) +
  theme(legend.position="right",
        plot.title = element_text(hjust = 0.5))
```

Judging by the logarithmic scale, no surprise removing rows and median replenishment are definetely the fastest methods, with removing rows being several times faster. Looking at more advanced ones, definetely *VIM* rules - probably usually something like 5-10 times faster than *mice* and *missForest*. These last two are quite slow, with the missforest appearing to be slightly faster.

What do the numbers say?

```{r imputation_times_measures, cache = TRUE}
means <- rep(NA, nimputations)
medians <- rep(NA, nimputations)
stDeviations <- rep(NA, nimputations)

for(i in (1:nimputations)){
  means[i] <- mean(imputation_times[imputation_names[i]][!is.na(imputation_times[imputation_names[i]])])
  means[i] <- round(means[i], 3)
  medians[i] <- median(imputation_times[imputation_names[i]][!is.na(imputation_times[imputation_names[i]])])
  medians[i] <- round(medians[i], 3)
  stDeviations[i] <- sd(imputation_times[imputation_names[i]][!is.na(imputation_times[imputation_names[i]])])
  stDeviations[i] <- round(stDeviations[i], 3)
}

imputation_times_measures <- data.frame(imputation_names, means, medians, stDeviations)
knitr::kable(imputation_times_measures)
```

Taking into acount all the imputations which have been implemented, we can clearly see that two easy ones really stand out. However, we cannot fully compare the other three - due to the fact that missForest failed on three data sets, and mice - up to seven.

Considering such a large spread of data size, it is very interesting difference between the median and the average for VIM - the first is almost seven minutes, the second - barely one and a half seconds. In general, however, it is certainly much longer than methods for removing incomplete rows and filling with median.

Let's count these four measures again, but only for those datasats for which all imputations were successful - these are ids 27, 38, 55, 56, 944, 188  (dataset with id 4 softened on removing rows containing any missing items, because each of its poems had some missing items).

```{r imputation_times_measures_little, cache = TRUE}
dataset_ids_fullimp <- c(2, 5, 8, 11, 13, 14)
ndatasets_fullimp <- length(dataset_ids_fullimp)

means <- rep(NA, nimputations)
medians <- rep(NA, nimputations)
stDeviations <- rep(NA, nimputations)

for(i in 1:nimputations){
  means[i] <- mean(imputation_times[imputation_names[i]][dataset_ids_fullimp,])
  means[i] <- round(means[i], 3)
  medians[i] <- median(imputation_times[imputation_names[i]][dataset_ids_fullimp,])
  medians[i] <- round(medians[i], 3)
  stDeviations[i] <- sd(imputation_times[imputation_names[i]][dataset_ids_fullimp,])
  stDeviations[i] <- round(stDeviations[i], 3)
}

imputation_times_measures_fullimp <- data.frame(imputation_names, means, medians, stDeviations)
knitr::kable(imputation_times_measures_fullimp)
```

Now we have better results to compare. Taking into account the mean time, definetely *missForest* is the slowest, but also its standard deviaton is incomparably huge - this is probably due to the fact that for smaller sets it is doing well, but due to its complexity, its slowdown can be seen for very large datasets. As shown in the chart, *VIM* is definitely better for quick calculations than *missForest* and *mice*, and considering the median, *mice* is comparable to *missForest* - so you can expect that for small data sets there is not much difference between them, and a lot of time we definitely need to devote to these larger data frames.

###########################################3

facet?
x - id zbioru, y - czas imputacji, z - procent braków
kategoryczny, ciągły, ciągły

facet?
x - id zbioru, y - czas imputacji, z - wielkość zbioru
kategoryczny, ciągły, ciągły

## Comparing modelling times

TODO: to co wyżej, ale na czasach modelowania. Pewnie wniosek, że czas modelowania mega podobny

## Best measures

TODO:
1. weź w CZTERY ramki danych cztery miary po zbiorach (x - id zbioru, y - wynik miary)
2. Z pomocą grid_arrange() połącz cztery ggploty z wykresami liniowymi
3. Wysnuj wnioski o najlepszych i najgorszych miarach na podstawie obrazka

4. Wylicz średnie i wariancje dla wszystkich analizowanych zbiorów dla czterech miar
5. Zrób prostego ggplota średnich, a imputacji
6. Zrób prostego ggplota wariancji, a imputacji
7. Wysnuj wnioski z powyższych dwóch wykresów na podstawie wyliczonych prostych miar

## imputation time and its results

1. Wykres po zbiorach: średnich czasów imputacji a średnich wyników miar.
2. Wnioski z tego wykresu. Czy są imputacji i lepsze i szybsze? Czy są i wolniejsze i gorsze? PYTANIE: jaki wykres, w jakiej formie?

## Conclusion

TODO: 1. Podsumowanie i wnioski

OPCJONALNIE: porównanie wielkości zbiorów a czasów imputacji